\section{Background}
\label{sec:background}

We categorize prior work into \emph{inference benchmarks} that measure serving performance and \emph{inference simulators} that predict it. Table~\ref{tab:comparison} provides a unified feature comparison.

\paragraph{Inference benchmarks.}
MLPerf Inference~\cite{mlperf} provides standardized throughput and latency measurements with SLO enforcement but uses fixed reference datasets and does not support agentic or multi-turn workloads.
SGLang~\cite{sglang} includes a benchmark suite with ShareGPT traces but focuses on single-turn chat.
InferenceX provides a public dashboard of hardware comparisons but uses random tokens with arbitrary input/output lengths that do not reflect real-world agentic workloads.
ML.ENERGY~\cite{mlenergy} benchmarks energy consumption across 46 models but does not measure latency or provide kernel-level analysis.
None of these tools provide per-kernel profiling or latency prediction capabilities.

\paragraph{Inference simulators.}
LLMCompass~\cite{llmcompass} uses analytical tile-by-tile simulation for hardware design exploration, achieving 4.1\% end-to-end error but does not model serving features (continuous batching, CUDA graphs) or validate against hardware counters.
Vidur~\cite{vidur} uses Random Forest interpolation with profiling data, supporting five scheduling policies with ${<}5\%$ error at 85\% capacity, but lacks per-kernel validation and MoE support.
LLMServingSim~v2~\cite{llmservingsim} achieves 0.97--1.54\% error with runtime-loop simulation and supports MoE, but uses lookup tables rather than generalizable ML and does not validate against ncu hardware counters.
DistServe~\cite{distserve} models SLO attainment for prefill-decode disaggregation with ${<}2\%$ error using queueing theory but only supports dense architectures.
Frontier~\cite{frontier} uses per-operator Random Forest prediction for MoE and disaggregated systems but reports 19--23\% end-to-end error.
No existing simulator validates per-kernel predictions against ncu hardware counter ground truth or links kernel-level resource utilization to serving-level performance.

\definecolor{okgreen}{RGB}{0,140,54}
\definecolor{nored}{RGB}{200,30,30}
\definecolor{partialorange}{RGB}{210,125,0}
\newcommand{\cmark}{\textcolor{okgreen}{\checkmark}}
\newcommand{\xmark}{\textcolor{nored}{\texttimes}}
\newcommand{\pmark}{\textcolor{partialorange}{$\sim$}}
\begin{table}[t]
\caption{Feature comparison of representative LLM inference simulators and benchmarks against our work. \cmark\ denotes full support, \xmark\ no support, and \pmark\ partial support. Reported accuracy is the end-to-end latency error stated by each system on its own validation setup (lower is better; ``---'' indicates the system does not predict latency). To our knowledge, ours is the first framework to unify agentic trace benchmarking with per-kernel roofline validation against hardware counters.}
\label{tab:comparison}
\centering
\small
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.1}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{l ccccccc}
\toprule
\textbf{Feature} & \textbf{LLMCompass} & \textbf{Vidur} & \textbf{LLMServ.\,v2} & \textbf{DistServe} & \textbf{MLPerf} & \textbf{ML.ENERGY} & \textbf{Ours} \\
 & \scriptsize ISCA'24 & \scriptsize MLSys'24 & \scriptsize CAL'25 & \scriptsize OSDI'24 & \scriptsize MLSys & \scriptsize NeurIPS'25 & \\
\midrule
\multicolumn{8}{l}{\emph{Prediction \& measurement}} \\
\quad Predicts latency                & \cmark & \cmark & \cmark & \cmark & \xmark & \xmark & \cmark \\
\quad Measures real serving           & \xmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
\quad Per-kernel \texttt{ncu} validation & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\quad Agentic workload profiles       & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\quad Multi-turn benchmarking         & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\midrule
\multicolumn{8}{l}{\emph{Model architecture}} \\
\quad Dense transformer               & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
\quad Mixture-of-experts              & \pmark & \xmark & \cmark & \xmark & \xmark & \cmark & \cmark \\
\quad Grouped-query attention         & \xmark & \cmark & \pmark & \xmark & \xmark & \cmark & \cmark \\
\midrule
\multicolumn{8}{l}{\emph{Serving features}} \\
\quad Tensor parallelism              & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
\quad Continuous batching             & \xmark & \cmark & \cmark & \cmark & \cmark & \cmark & \xmark \\
\quad CUDA graphs                     & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \pmark \\
\midrule
\multicolumn{8}{l}{\emph{Analysis}} \\
\quad Roofline analysis               & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\quad Kernel resource utilization     & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\quad Saturation curves               & \xmark & \xmark & \xmark & \xmark & \xmark & \xmark & \textbf{\cmark} \\
\midrule
\multicolumn{8}{l}{\emph{Validation \& availability}} \\
\quad Reported latency error          & 4.1\% & ${<}5\%$ & 0.97\% & ${<}2\%$ & --- & --- & \textbf{2\%} \\
\quad Public profiling dataset        & \xmark & \xmark & \xmark & \xmark & \xmark & \cmark & \textbf{\cmark} \\
\quad Open source                     & \pmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}
