\section{Introduction}
\label{sec:intro}

Current popular LLM inference benchmarks often use traces which fail to properly represent real-life workloads. These include short single-turn chat exchanges, synthetic random-token stress tests, or fixed-length request distribution drawn from datasets like ShareGPT~\citep{sharegpt}. MLPerf Inference~\citep{mlperf,mlperf_v51} evaluates summarisation, chat, and long-context reasoning; its LLM workloads range from 1{,}024-token Q\&A (Llama~2~70B on OpenOrca) to 128K-token document tasks (Llama~3.1~405B), but remain single-turn with no multi-step tool use. The benchmarks bundled with vLLM~\citep{vllm} default to ShareGPT prompts whose input lengths cluster around 1{,}024 tokens; SGLang~\citep{sglang} similarly defaults to ShareGPT for serving evaluation.

Using single-turn datasets to create benchmarks leads to misleading information on actual GPU utilisation on real workloads. For example, ShareGPT has a typical ISL/OSL ratio of ... , with token amount ranging in the thousands. In contrast, a coding agent processing a SWE-Bench (cite?) task sends a reposity-scale context of tens of thousands of tokens and receives a short code patch, giving a typical ISL/OSL ratio exceeding 20:1, a context length ratio far outside the range of any existing benchmark evaluation. This is not simply a quantitative difference but a qualitative regime shift: chat workloads a predominantly decode-dominated and memory-bandwidth-bound at typical serving concurrencies, while agentic workloads shift time toward compute-bound prefill GEMMs and generate sustained KV cache pressure from growing multi-turn context.

The practical consequence is severe. On an H100 GPU, Llama-3.1-8B sustains XXX output tokens/s under a chat-short workload but only XXX tokens/s under a coding-agent workload at matched concurrency---a X.X$\times$ gap. Moreover, existing benchmarks rarely sweep concurrency to saturation; most report a single operating point or a narrow range, missing the throughput ceiling and the latency degradation curve that determine real-world capacity.

We present \textsc{AgentServe}, a benchmark suite and analysis framework for LLM inference under agentic workloads. Our contributions are:

-- Anything below is not properly proofread and is AI generated --

\begin{enumerate}
    \item \textbf{Saturation benchmarks on agentic workloads.} We benchmark 8 models across 3 families (Llama-3, Qwen, GPT-OSS; dense and MoE, 8B--120B parameters) on A100 and H100 GPUs using both vLLM and SGLang, sweeping concurrency from 1 to 320 with continuous batching (Section~\ref{sec:results}). We produce 2{,}400+ result configurations. Agentic workloads saturate at 2--4$\times$ lower throughput than chat workloads across all model--hardware combinations and exhibit steeper TTFT degradation under load.

    \item \textbf{Per-kernel roofline analysis.} Using Nsight Compute hardware counters, we profile 2{,}900+ CUDA kernels across 48 model configurations (Section~\ref{sec:kernel_roofline}). The throughput gap traces to a regime shift: chat requests spend GPU time in memory-bound decode GEMMs (arithmetic intensity ${\approx}$1, near peak HBM bandwidth), while agentic requests shift time to compute-bound prefill GEMMs (arithmetic intensity $>$500, near peak TFLOPS). MoE models generate 3.7$\times$ more kernel invocations than dense models due to per-expert dispatch. These per-kernel data inform choices for prefill-decode disaggregation (PDD) and attention-FFN disaggregation (AFD) in agentic serving configurations.

    \item \textbf{Per-operator latency predictors.} We train XGBoost models on the collected kernel profiles, enabling latency estimation for new architectures without re-running GPU experiments (Section~\ref{sec:predictor}). The hybrid predictor achieves 0.81--1.00$\times$ accuracy on decode latency and selects automatically between ML and analytical modes per inference phase.
\end{enumerate}

We release all benchmark results, ncu kernel profiles, workload traces, trained predictors, and the benchmarking tool under an open-source license.\footnote{Dataset and code: \url{[anonymized for review]}.}