\section{Methodology}
\label{sec:method}

\subsection{Workload Taxonomy}
\label{sec:taxonomy}

We organize inference workloads into three tiers (Table~\ref{tab:profiles}). Tier~1 (agentic) and the multi-turn variants of Tier~2 are the primary contribution; Tier~2 single-turn and Tier~3 act as controls.

\begin{itemize}
    \item \textbf{Tier~1: Agentic traces.} Real request logs from SWE-Bench~\citep{swebench} coding agents, TerminalBench CLI agents, and OSWorld computer-use agents. SWE-Bench sessions span 13--320 turns (median 85, $n{=}165$); TerminalBench sessions span 2--876 turns (median 61, $n{=}267$); OSWorld covers 60 sessions of WebArena traces. Each multi-turn variant (\texttt{short}/\texttt{medium}/\texttt{long}/\texttt{xl}) is bucketed by turn count and capped at a maximum context window of 32K, 65K, or 131K tokens. We also include a single-turn \texttt{coding-agent} profile (per-request ISL$\approx$17K, OSL$\approx$800) as the agentic single-turn baseline.

    \item \textbf{Tier~2: Chat baselines.} ShareGPT-derived conversations at three single-turn scales (chat-short/medium/long: ISL 500/2K/8K, OSL 300/1K/2K) and three multi-turn scales (3--20 turns, max context 8K--32K). Source: \texttt{Aeala/ShareGPT\_Vicuna\_unfiltered}; conversations are filtered by an approximate token count (word\_count $\times 1.35$) so the listed ISL acts as an upper-bound filter rather than an exact per-request length.

    \item \textbf{Tier~3: Synthetic stress tests.} Random-token workloads isolating specific bottlenecks: \texttt{decode-heavy} (ISL=256, OSL=4{,}096) for memory-bandwidth stress, \texttt{prefill-heavy} (ISL=8{,}192, OSL=256) for compute stress, and \texttt{random-1k} (ISL=1{,}024, OSL=1{,}024).
\end{itemize}

\paragraph{Single-turn vs.\ multi-turn semantics.} For single-turn profiles the ISL column is the (approximate) per-request input length. For multi-turn profiles, the ISL column is the \emph{maximum context window cap} for the session: turn~1 starts with the system prompt plus the first user message ($\sim$2K tokens), and the context grows by user/assistant pairs each turn until it hits the cap or the session ends.

\begin{table}[t]
\caption{Workload profiles. For single-turn profiles, ISL/OSL are approximate per-request input/output sequence lengths. For multi-turn profiles, ISL is the \emph{maximum context window cap} (context grows across turns) and OSL is the per-turn output cap. The \texttt{coding-agent} single-turn ratio (21$\times$) is 12$\times$ that of \texttt{chat-short}; the longest multi-turn profiles reach 65--131$\times$ at session end.}
\label{tab:profiles}
\centering
\small
\begin{tabular}{llrrrrrr}
\toprule
\textbf{Tier} & \textbf{Profile} & \textbf{ISL} & \textbf{OSL} & \textbf{ISL/OSL} & \textbf{Turns} & \textbf{Sess.} & \textbf{Source} \\
\midrule
1 & coding-agent                     & 17{,}000  & 800   & 21.3 & 1       & 200 & SWE-Bench \\
1 & swebench-multiturn-short         & 32{,}768  & 2{,}000 & 16.4 & 13--30  & 100 & SWE-Bench \\
1 & swebench-multiturn-medium        & 65{,}536  & 2{,}000 & 32.8 & 30--80  & 100 & SWE-Bench \\
1 & swebench-multiturn-long          & 131{,}072 & 2{,}000 & 65.5 & 80--150 & 50  & SWE-Bench \\
1 & swebench-multiturn-xl            & 131{,}072 & 2{,}000 & 65.5 & 150--400 & 30 & SWE-Bench \\
1 & terminalbench-multiturn-short    & 32{,}768  & 2{,}000 & 16.4 & 2--20   & 100 & TerminalBench \\
1 & terminalbench-multiturn-medium   & 65{,}536  & 2{,}000 & 32.8 & 20--60  & 100 & TerminalBench \\
1 & terminalbench-multiturn-long     & 131{,}072 & 2{,}000 & 65.5 & 60--150 & 50  & TerminalBench \\
1 & terminalbench-multiturn-xl       & 131{,}072 & 2{,}000 & 65.5 & 150--1000 & 30 & TerminalBench \\
1 & osworld-multiturn-short          & 32{,}768  & 500   & 65.5 & 2--10   & 50  & OSWorld \\
1 & osworld-multiturn-medium         & 65{,}536  & 500   & 131.1 & 10--20 & 30  & OSWorld \\
\midrule
2 & chat-short                       & 500      & 300   & 1.7  & 1       & 200 & ShareGPT \\
2 & chat-medium                      & 2{,}000  & 1{,}000 & 2.0 & 1       & 200 & ShareGPT \\
2 & chat-long                        & 8{,}000  & 2{,}000 & 4.0 & 1       & 200 & ShareGPT \\
2 & chat-multiturn-short             & 8{,}192  & 1{,}000 & 8.2 & 3--5    & 200 & ShareGPT \\
2 & chat-multiturn-medium            & 16{,}384 & 1{,}500 & 10.9 & 5--10  & 100 & ShareGPT \\
2 & chat-multiturn-long              & 32{,}768 & 2{,}000 & 16.4 & 10--20 & 50  & ShareGPT \\
\midrule
3 & decode-heavy                     & 256      & 4{,}096 & 0.06 & 1     & 200 & Random \\
3 & prefill-heavy                    & 8{,}192  & 256   & 32.0 & 1       & 200 & Random \\
3 & random-1k                        & 1{,}024  & 1{,}024 & 1.0 & 1       & 200 & Random \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Benchmarking Framework}
\label{sec:framework}

We build on an open-source asynchronous benchmarking tool that sends concurrent requests to OpenAI-compatible serving endpoints. For each (model, profile, concurrency) configuration, we record per-request metrics from the streamed response: time-to-first-token (TTFT), inter-token latency (ITL), time-per-output-token (TPOT), end-to-end latency, and input/output token counts. We report median and p90 percentiles for latency metrics and aggregate throughput in output tokens per second. Three requests are issued as warmup and discarded before measurement.

\paragraph{Arrival pattern and concurrency.} The framework uses a \emph{closed-loop, steady-concurrency} arrival pattern implemented as an \texttt{asyncio.Semaphore(concurrency)}: all $N$ requests are submitted at $t{=}0$ but at most \texttt{concurrency} are in flight at any moment, with the next request firing immediately upon completion of an in-flight one. Unlike open-loop \texttt{--request-rate inf} runs (which inflate TTFT with queue-wait time), our TTFT reflects actual scheduler behavior at a stable in-flight load. We sweep concurrency over $\{1, 10, 20, 40, 80, 120, 160, 200, 256, 320\}$ on A100 and H100, and over $\{1, 10, 20, 40, 80\}$ on RTX~3090 (memory-limited). The number of measured requests scales with concurrency: $N{=}50$ at conc.\ 1, $N{=}200$ at conc.\ 5--80, and $N{=}\max(200, 2\cdot\text{conc.})$ at conc.\ ${\geq}120$, ensuring the system reaches steady state at every operating point.

\paragraph{Statistical caveat.} Each (model, profile, concurrency) cell is run once with the $N$ requests above; percentiles are computed across those requests rather than across multiple repeats. We do not report p99 because $N{=}50$ at concurrency~1 makes p99 a single-sample estimate; p50 and p90 are the headline statistics throughout.

\subsection{Kernel-Level Profiling}
\label{sec:ncu_method}

To explain \emph{why} different workload profiles produce different throughput characteristics, we profile the CUDA kernels executed during model forward passes using NVIDIA Nsight Compute (ncu, v2024.3.2) with the \texttt{--set full} flag, which collects the full counter set per kernel:
\begin{itemize}
    \item \textbf{Compute throughput}: SM active cycles, tensor core utilization, achieved TFLOPS
    \item \textbf{Memory throughput}: DRAM bytes read/written, L2 cache hit rate, achieved HBM bandwidth
    \item \textbf{Kernel duration}: wall-clock execution time in microseconds
\end{itemize}

We profile two complementary datasets used for predictor training (Section~\ref{sec:perop_method}) and roofline analysis (Section~\ref{sec:roofline}):
\begin{itemize}
    \item \textbf{2-layer model forward passes.} Llama-3.1-8B (417 kernels), Mixtral-8$\times$7B (1{,}525 kernels), and two held-out unseen models (Qwen-2.5-72B, Llama-3.3-70B; 421 kernels each) at batch size~1 and sequence length~128, capturing the full kernel mix (flash attention, RMSNorm, SiLU, residual additions, MoE expert routing).
    \item \textbf{GEMM shape sweeps.} 56 isolated GEMM shapes spanning $M \in \{1,2,4,\ldots,8192\}$ and $N,K \in \{4096,8192,14336\}$ (372 kernels), plus a wider hidden-size sweep (451 kernels) covering hidden $\in \{2048,\ldots,28672\}$.
\end{itemize}
This yields 2{,}811 profiled kernels in total, with 842 reserved as held-out validation.

\paragraph{Why 2 layers suffice.} The kernel mix is identical across all transformer blocks: each layer issues the same operators with the same tensor shapes. Only the KV-cache size grows with layer index, but our ncu profiles capture prefill at sequence length~128 with no accumulated cache. We verify experimentally that summing 2-layer kernel times and scaling by $n_{\text{layers}}/2$ matches a full-model forward pass within $+0.1$--$0.2\%$ error.

\paragraph{Roofline classification.}
From the ncu data we compute each kernel's \emph{arithmetic intensity} (AI = FLOPs / bytes transferred) and plot it against achieved throughput. We use the A100's \emph{dense} bf16 tensor core peak: 312 TFLOPS / 2{,}039 GB/s HBM2e $\Rightarrow$ ridge AI $= 153$ FLOPs/byte. Kernels with AI $> 153$ are compute-bound; AI $< 153$ are memory-bound. (At our profiling configuration of bs=1, seq=128, all profiled kernels lie in the memory-bound region; the AI distribution shifts right as $M$ grows under continuous batching, which is exactly the regime studied in Section~\ref{sec:roofline}.)

\subsection{Per-Operator and Per-Kernel Latency Predictors}
\label{sec:perop_method}

To enable latency estimation without re-running GPU experiments for every new model or shape, we train two complementary XGBoost regressors operating at different granularities.

\paragraph{Level~1: per-kernel predictor.} Trained on the 2{,}811 ncu-profiled kernels with a log-transformed latency target. Features (41 total) combine raw ncu counters and derived quantities: launch configuration (block/grid size, registers, shared memory, waves, occupancy, thread count); memory traffic (DRAM read/write bytes, L1TEX/LTS hit rate); throughput (SM\%, DRAM\%); cycle counts (active, elapsed); plus arithmetic intensity, total/log DRAM bytes, active-cycle ratio, bytes-per-thread, write/read ratio, bytes-per-cycle, grid volume, and a one-hot encoding of 10 kernel families (gemm, flash\_attn, reduce, elementwise, copy, compact, topk, cast, index, random). Hyperparameters: \texttt{n\_estimators=500}, \texttt{max\_depth=6}, \texttt{lr=0.03}.

\paragraph{Level~2: per-operator predictor.} Trained on 13{,}144 CUDA-event timing samples collected from end-to-end forward passes of the 6 evaluated models (Llama-3.1-8B/70B, Mixtral-8$\times$7B, Qwen-2.5-72B, gpt-oss-20b/120b). To avoid pipeline serialization, the timing pre-hook does not call \texttt{cuda.synchronize}; only the post-hook does. The model uses 26 analytical features---\{tok, log\_tok, FLOPs, log\_FLOPs, weight\_bytes, log\_wt, AI, log\_AI, op-type one-hot (attn/ffn/norm), $d_\text{model}$, $d_\text{ffn}$, num\_experts, top-$k$, $d^2$, $d \cdot d_\text{ffn}$, effective\_tok, active\_experts, total\_weight, batch\_size, seq\_len, log\_bs, log\_seq, attn\_quadratic, log\_attn\_quad\}---predicting raw per-op latency in microseconds. Hyperparameters: \texttt{n\_estimators=200}, \texttt{max\_depth=6}, \texttt{lr=0.1}.

\paragraph{Validation protocol.} We use leave-one-source-out (per-kernel) and leave-one-model-out (per-op) cross-validation, plus held-out evaluation on two models excluded from training (Qwen-2.5-72B and Llama-3.3-70B). Accuracy results are reported in Section~\ref{sec:results}.

\subsection{Multi-Node Serving}
\label{sec:multinode}

TBD.

\subsection{Experimental Setup}
\label{sec:setup}

\paragraph{Hardware.}
We evaluate on three GPU platforms: (1)~8$\times$ NVIDIA A100-SXM4-40GB on-premise (NVSwitch fabric, \texttt{nvidia-fabricmanager}); (2)~2--4$\times$ NVIDIA H100-SXM5-80GB cloud (RunPod, NVLink); and (3)~8$\times$ NVIDIA RTX~3090 24GB on-premise (PCIe, no NVLink). The A100 provides 312 TFLOPS bf16 / 2{,}039 GB/s HBM2e; the H100 provides 989 TFLOPS bf16 / 3{,}352 GB/s HBM3. Each benchmark run uses the subset of GPUs required by the TP degree (e.g., TP=4 uses GPUs 0--3) and runs in isolation (no co-location).

\paragraph{Software.}
vLLM v0.19.0; SGLang as a secondary comparison on H100; CUDA~12.6 (A100, ncu profiling) / 13.0 (RTX~3090); NVIDIA driver 580.126.09 on the 3090 cluster; PyTorch~2.10; XGBoost~3.2.0; ncu (Nsight Compute) v2024.3.2; Python~3.12 (conda environment).

\paragraph{Serving flags.}
On A100 we launch vLLM with \texttt{--enforce-eager --disable-custom-all-reduce --no-enable-log-requests --dtype bfloat16 --gpu-memory-utilization 0.95}; \texttt{--enforce-eager} is required because CUDA-graph capture OOMs at our memory budget, and the custom all-reduce kernel was unstable on A100-SXM4. \texttt{--max-model-len} is set to 4096 for gpt-oss-20b and 2048 for the larger models. On H100 we use vLLM defaults (CUDA graphs enabled, \texttt{--dtype bfloat16}). Prefix caching is disabled; chunked prefill is on by default in vLLM~0.19. All models are served in bfloat16 except gpt-oss-20b and gpt-oss-120b, which use the released MXFP4 weights.

\paragraph{Tokenization.}
Requests are sent as chat messages over the OpenAI-compatible API; tokenization and chat templating are handled by the serving engine using each model's native tokenizer. For ShareGPT and agent traces, the ISL value listed in Table~\ref{tab:profiles} is an \emph{upper-bound filter} (word\_count $\times 1.35$), not an exact per-request token count. Random/synthetic profiles sample tokens uniformly from the vocabulary.

\paragraph{Models.}
Table~\ref{tab:models} summarizes the 6 model architectures evaluated, spanning dense transformers (Llama-3.1-8B/70B, Qwen-2.5-72B) and mixture-of-experts (Mixtral-8$\times$7B; gpt-oss-20b with 16 experts top-4; gpt-oss-120b with 128 experts top-4). Each model is run at the TP degrees feasible for its memory footprint on each platform (Table~\ref{tab:tp_configs}).

\begin{table}[t]
\caption{Model configurations evaluated. Dense models use grouped-query attention (GQA); MoE models use top-$k$ expert routing.}
\label{tab:models}
\centering
\small
\begin{tabular}{lcrrrrrrcc}
\toprule
\textbf{Model} & \textbf{Type} & \textbf{Params} & \textbf{$d_\text{model}$} & \textbf{Layers} & \textbf{Heads} & \textbf{KV} & \textbf{$d_\text{ffn}$} & \textbf{Experts} & \textbf{top-$k$} \\
\midrule
Llama-3.1-8B        & Dense & 8B   & 4{,}096 & 32 & 32 & 8 & 14{,}336 & ---  & --- \\
Llama-3.1-70B       & Dense & 70B  & 8{,}192 & 80 & 64 & 8 & 28{,}672 & ---  & --- \\
Qwen-2.5-72B        & Dense & 72B  & 8{,}192 & 80 & 64 & 8 & 29{,}568 & ---  & --- \\
Mixtral-8$\times$7B & MoE   & 47B  & 4{,}096 & 32 & 32 & 8 & 14{,}336 & 8    & 2 \\
gpt-oss-20b         & MoE   & 20B  & 6{,}144 & 52 & 48 & 8 & 16{,}384 & 16   & 4 \\
gpt-oss-120b        & MoE   & 120B & 6{,}144 & 96 & 48 & 8 & 16{,}384 & 128  & 4 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\caption{Tensor-parallel configurations actually run on each platform. gpt-oss models use MXFP4 weights (released format); all others are bfloat16. Llama-70B/Qwen-72B require TP=4 on A100-40GB because the bf16 weights ($\sim$140\,GB) do not fit on 2$\times$40\,GB.}
\label{tab:tp_configs}
\centering
\small
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{A100 (8$\times$40GB)} & \textbf{H100 (2$\times$80GB)} & \textbf{RTX 3090 (8$\times$24GB)} \\
\midrule
Llama-3.1-8B        & TP=1            & TP=1            & TP=1, 2 \\
Llama-3.1-70B       & TP=4            & TP=2, TP=4      & --- \\
Qwen-2.5-72B        & TP=4            & TP=2, TP=4      & --- \\
Mixtral-8$\times$7B & TP=4            & TP=2            & --- \\
gpt-oss-20b         & TP=2 (MXFP4)    & TP=1, TP=2      & TP=1 \\
gpt-oss-120b        & TP=8 (MXFP4)    & TP=2 (MXFP4)    & --- \\
\bottomrule
\end{tabular}
\end{table}
